### 2017-01-29 manuscript.rmd
typo, changed unit years to My

In order to better understand why our phylogenetic analysis seems to support a different conclusion (*i.e.*, no support for the ortholog conjecture) than the published analysis of KMRR [@KryuchkovaMostacci:2016iw] (*i.e.*, strong support for the ortholog conjecture), we first checked to make sure we could reproduce their result based on pairwise analyses. This is important since we are only looking at at a subset of the data they considered, the Brawand *et al.* 2011 [@Brawand:2011du] dataset for gene trees that could be successfully time calibrated. We can replicate their result - we find a higher Tau correlation between orthologs (R=`r round( ortholog_r, 3) `) then between paralogs (R=`r round( paralog_r, 3)`).

Why is it that pairwise methods and phylogenetic methods lead to opposite conclusions? The reason is that they describe different things. Pairwise comparisons describe contemporary patterns, while phylogenetic methods infer historical processes. There need not be a different process of evolution following speciation and duplication for paralogs to be more different than orthologs. This could be due to the structure of the gene phylogenies. If paralogs tend to be more distantly related to each other than orthologs, then there would be more time for differences to accumulate even if the rate of change is the same between the two. This is, in fact, the case for these data. While the mean distance between orthologs is `r round( ortholog_distance_mean, 1 )` years, the mean distance between paralogs is `r round( paralog_distance_mean, 1 )` years. As described above, this is because the oldest speciation event is by definition the most recent common ancestor of the the species included in the study, but many gene families underwent duplication before this time.

>>>>

In order to better understand why our phylogenetic analysis seems to support a different conclusion (*i.e.*, no support for the ortholog conjecture) than the published analysis of KMRR [@KryuchkovaMostacci:2016iw] (*i.e.*, strong support for the ortholog conjecture), we first checked to make sure we could reproduce their result based on pairwise analyses. This is important since we are only looking at at a subset of the data they considered, the Brawand *et al.* 2011 [@Brawand:2011du] dataset for gene trees that could be successfully time calibrated. We can replicate their result - we find a higher Tau correlation between orthologs (R=`r round( ortholog_r, 3) `) than between paralogs (R=`r round( paralog_r, 3)`).

Why is it that pairwise methods and phylogenetic methods lead to opposite conclusions? The reason is that they describe different things. Pairwise comparisons describe contemporary patterns, while phylogenetic methods infer historical processes. There need not be a different process of evolution following speciation and duplication for paralogs to be more different than orthologs. This could be due to the structure of the gene phylogenies. If paralogs tend to be more distantly related to each other than orthologs, then there would be more time for differences to accumulate even if the rate of change is the same between the two. This is, in fact, the case for these data. While the mean distance between orthologs is `r round( ortholog_distance_mean, 1 )` My, the mean distance between paralogs is `r round( paralog_distance_mean, 1 )` My. As described above, this is because the oldest speciation event is by definition the most recent common ancestor of the the species included in the study, but many gene families underwent duplication before this time.
